---
title: "NYC AirBnB"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

You can find details about how I obtained the NYC Airbnb dataset <a href="airbnb_nyc_dataset.html" target="_blank">here</a> 

Let's set some global options
```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

library(tidyverse)
library(viridis)
```

Let's clean the data a bit

```{r}
nyc_airbnb = read_csv("data/nyc_airbnb.csv") %>%
  select(-X1) %>% 
  mutate(stars = review_scores_location/2) %>% 
  rename(borough = neighbourhood_group)
  
```

Now that we have our data and libraries ready. Let the exploration begin! 

## Some interesting questions about AirBnb! 

* Does rating vary by neighborhood, room type, or both?</br>
* How is price related to other variables? <br>
* Where are rentals located? <br>
* What's the most and least expensive area? <br>
* Which area has the highest density of listings?<br>
* What are some host characteristics? How many listings? What's associated with ratings?<br>

```{r}
nyc_airbnb %>% 
  count(room_type) %>% knitr::kable()

nyc_airbnb %>% count(borough) %>%
  arrange(desc(n)) %>% 
  knitr::kable() 
```

#### Does rating vary by neighborhood, room type, or both?

```{r}
nyc_airbnb %>%
 group_by(neighbourhood) %>%
 summarize(median = median(stars, na.rm = TRUE), 
           mean = mean(stars, na.rm = TRUE),
           number_of_listings = n()) %>% 
  knitr::kable()

nyc_airbnb %>%
  group_by(room_type) %>%
  summarize(median_stars = median(stars, na.rm = TRUE), 
            average_stars = mean(stars, na.rm = TRUE),
           number_of_listings = n())
```

```{r}
ggplot(nyc_airbnb, aes(x = stars), color = boro) +
 geom_histogram() +
 facet_grid(~room_type)
```

```{r}
nyc_airbnb %>%
  group_by(borough, room_type) %>%
  summarise(
  mean_stars = mean(stars, na.rm = TRUE)) %>% 
  spread(key = room_type, value = mean_stars)
```

#### Where are rentals located?

```{r}
nyc_airbnb %>% 
  filter(borough == "Manhattan", price < 1000) %>% 
  ggplot(aes(x = longitude, y = latitude, color = price)) + 
  geom_point(alpha = .2) + 
  facet_grid(~room_type) + 
  coord_cartesian() + 
  scale_color_viridis()
```

Entire homes are more expensive than other room types, and are more clustered in southern Manhattan. 

#### Expensive areas

```{r}
nyc_airbnb %>%
 select(borough, neighbourhood, price) %>%
 filter(borough == "Manhattan") %>%
 group_by(neighbourhood) %>%
 summarize(mean_price = mean(price, na.rm = TRUE),
           median_price = median(price, na.rm = TRUE)) %>% 
  top_n(5, median_price) %>%
 arrange(desc(median_price))

nyc_airbnb %>%
 select(borough, neighbourhood, price) %>%
 filter(borough == "Brooklyn") %>%
 group_by(neighbourhood) %>%
 summarize(mean_price = mean(price, na.rm = TRUE),
           median_price = median(price, na.rm = TRUE)) %>% 
  top_n(5, median_price) %>%
 arrange(desc(median_price))

nyc_airbnb %>%
 select(borough, neighbourhood, price) %>%
 filter(borough == "Manhattan") %>%
 group_by(neighbourhood) %>%
 summarize(mean_price = mean(price, na.rm = TRUE),
           median_price = median(price, na.rm = TRUE)) %>% 
  top_n(-5, median_price) %>%
 arrange(median_price)

nyc_airbnb %>%
 select(borough, neighbourhood, price) %>%
 filter(borough == "Brooklyn") %>%
 group_by(neighbourhood) %>%
 summarize(mean_price = mean(price, na.rm = TRUE),
           median_price = median(price, na.rm = TRUE)) %>% 
  top_n(-5, median_price) %>%
 arrange(median_price)
```

