---
title: "AirBnB Dataset"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

The data was assessed from [Inside Airbnb](http://insideairbnb.com/get-the-data.html) on 3rd November, 2018. <br>

According to Inside Airbnb, they are "an independent, non-commercial set of tools and data that allows you to explore how Airbnb is really being used in cities around the world". You can learn more about Inside AirBnb [here](http://insideairbnb.com/index.html)<br>

The code below was used to create the final dataset for analysis

```{r}
library(tidyverse)

# First, uncompress and load the data. Then select id and review_scores_location variables

uncompress = gzfile(description = "data/listings.csv.gz", open = "rt")
df1 = read.csv(uncompress, header = T) %>% 
    select(id, review_scores_location)

# Read the summary for NYC
df2 <- read.csv("data/listings.csv", stringsAsFactors = FALSE) %>% 
  mutate(last_review = as.Date(last_review, format = "%Y-%m-%d"))

# Combine the two datasets
df <- inner_join(df1, df2, by = "id")

# Save the data
write.csv(df, file = "data/nyc_airbnb.csv")
```

The final dataset has 50,041 observations and 17 variables:

- `id`: listing id
- `review_scores_location`: 0-5 stars converted into a 0-10 scale
- `name`: listing name
- `host_id`: host id 
- `host_name`: host name
- `neighbourhood_group`: NYC borough
- `neighbourhood`: NYC neighborhood
- `latitude`: listing latitude
- `longitude`: listing longitude
- `room_type`: type of listing (Entire home/apt, Private room, Shared room)
- `price`: listing price
- `minimum_nights`: required minimum nights stay
- `number_of_reviews`: total number of reviews
- `last_review`: date of last review
- `reviews per month`: average number of reviews per month
- `calculated_host_listings_count`: total number of listings for this host
- `availability_365`: number of days listing is available out of 365

