---
title: "WH Gender Pay Gap"
---

---
title: "2017 White House Employee Salary Analysis - Gender pay gap"
author: "Amin Yakubu"
date: "10/6/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .9,
  out.width = "90%"
)
```

Loading tidyverse package and setting global options
```{r}
library(tidyverse)
library(knitr)
library(kableExtra)
library(plotly)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_bw() + theme(legend.position = "bottom"))
```


Reading and cleaning the dataset 
```{r}
wh_pay = read_csv("./data//white_house_2017_salaries_clean.csv") %>% 
  select(-(starts_with("X")))
```

In 2017, the White House spent a total of $ `r sum(wh_pay$SALARY)` on the salaries of `r nrow(wh_pay)` employees. 

Below is a table showing the top earner in the White House for 2017

```{r}
wh_pay %>% 
  top_n(1, SALARY) %>% 
  arrange(desc(SALARY)) %>% 
  knitr::kable() %>% 
  kable_styling(bootstrap_options = "striped", full_width = T)

```

### Distribution of top earnings
<br>

```{r}
salary = wh_pay %>% 
  filter(SALARY > 100000) %>% 
  group_by(GENDER) %>% 
  summarize(NUMBER = n()) %>% 
  ggplot(aes(x = GENDER, y = NUMBER, fill = GENDER)) + geom_col(width = .6) + viridis::scale_fill_viridis(discrete = TRUE)

ggplotly(salary)
```

</br>
Of the 153 staff earning more than $100,000 only 51 of them are female. 

```{r}
wh_pay %>% 
  filter(SALARY < 100000) %>% 
  group_by(GENDER) %>% 
  summarize(NUMBER = n()) %>% 
  knitr::kable() %>% 
  kable_styling(bootstrap_options = "striped", full_width = F, position = "float_right")
```

<p> <br> Of the remaining 224 that earn less than $100,000, 115 of them are female. This include Ivanka Trump and her husband, Jared Kusner, and Reed Cordish all of whom have a salary of 0. Gary Cohn, millionare former CEO of Goldman Sachs has a salary of 30,000. </br> </p>

```{r}
wh_pay %>%
  group_by(GENDER) %>%
  summarize(n = n()) %>% knitr::kable()

wh_pay %>% 
  group_by(TITLE, GENDER) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n)) %>% 
  knitr::kable()
```

```{r}
wh_pay %>% 
  group_by(GENDER) %>% 
  summarize(n = n(),
            mean = mean(SALARY),
            median = median(SALARY)) %>% knitr::kable()
```

```{r}
wh_pay %>%
  group_by(GENDER) %>%
  summarize(n = n(),
            mean = mean(SALARY),
            median = median(SALARY))  %>%
  ggplot(aes(x = GENDER, y = mean, fill = GENDER)) + 
  geom_col(width = 0.5) 
wh_pay %>%
  group_by(GENDER) %>%
  summarize(n = n(),
            mean = mean(SALARY),
            median = median(SALARY))  %>%
  ggplot(aes(x = GENDER, y = median, fill = GENDER)) + 
  geom_col(width = 0.5) 
ggplot(wh_pay, aes(x = GENDER, y = SALARY, fill = GENDER)) +
  geom_boxplot()
ggplot(wh_pay, aes(x = STATUS, fill = GENDER)) + 
  geom_bar(width = 0.5)
ggplot(wh_pay, aes(x = SALARY, fill = GENDER)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")
ggplot(wh_pay, aes(x = SALARY, fill = GENDER)) + 
  geom_histogram(position = "dodge") 
```
